Developing a preprint culture in biology

### 2012-09-26 ms.tex
change title
merging ethans branch
minor edits

\title{Open Preprints in Ecology \& Evolution}

@@

Yet, despite the success of the approach in other fields of inquiry, most
manuscripts in ecology and evolution are never submitted to any public archive.
In this article, we will first highlight the advantages of open preprints
servers for both scientists and publishers. We will also debunk a few
misconceptions, discuss the policies of major publishers in ecology an
evolution, and briefly review the most popular open preprint servers.

@@

\section{Preprints, Ecology \& Evolution}

While submitting to public preprints servers is still uncommon in ecology and
evolution, preprints are becoming more common in biological sciences. The
quantitative biology section in arXiv is experiencing faster growth in
submissions than any other field \cite{cal12}. Also, most scientific journals
are preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America recently changed its policy to allow
public preprints (REF). In our field, few scientific publications will not
consider a manuscript submitted to arXiv.  Still, many ecology \& evolution
journals adopt a ``by default'' hostile attitude towards preprints, mostly due
to the lack of clear policy of the publishers. As an example, Wiley-Blackwell,
which publishes some of the leading journal in the field, has no official policy
on the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

There are other ``cultural'' barriers to adoption of preprint servers by
ecologists and evolutionary biologists.  The most notable preprint server,
arXiv, is designed to work best with \TeX{}, a nearly universal document
preparation system among physicists and mathematicians.  Jackson
\cite{jackson2002preprints} further argues that \TeX{} introduced an ``open
source mindset'' to its users, who now freely share their research findings as
well as their software.  Many biologists instead use proprietary software to
prepare their research findings, and many ecology and evolution journals
officially prefer Microsoft Word documents as submissions.  The recent
discipline-wide adoption of free software packages such as R and associated
document generation systems \cite{xie2012}, as well as interest in open access
publishing, has coincided with recent rise in use of preprint servers by
biologists, supporting Jackson's claims. \cite{xie12}

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in mathematics and most branches of physics.
Physicist Paul Ginsparg originated arXiv in 1991 for theoretical high-energy
physicists to communicate preprints via email and ftp, and soon thereafter
adopted the newly created world-wide web\cite{jackson2002preprints}.  arXiv now
receives over 7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}).  arXiv divides its
submissions into subcategories of physics, mathematics, computer science,
quantitative biology, finance and statistics.  The quantitative biology category
includes subcategories for Populations and Evolution, Quantitative Methods and
other categories that may be of interest to ecologists and evolutionary
biologists.

@@

>>>>

\title{Developing a preprint culture in biology}

@@

Yet, despite the success of the approach in other fields of inquiry, most
manuscripts in biology are never submitted to any public archive.  In this
article, we will first highlight the advantages of open preprints servers for
both scientists and publishers. We will also debunk a few misconceptions,
discuss the policies of major publishers in biology, and briefly review the most
popular open preprint servers.

@@

\section{Preprints in biological sciences}

While submitting to public preprints servers is still uncommon in biology,
preprints are becoming more common in biological sciences. The quantitative
biology section in arXiv is experiencing faster growth in submissions than any
other field \cite{cal12}. Also, most scientific journals are preprint-friendly:
Nature, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the
journals from Elsevier and Springer. Very recently, the Ecological Society of
America recently changed its policy to allow public preprints (REF). In our
field, few scientific publications will not consider a manuscript submitted to
arXiv.  Still, many journals adopt a ``by default'' hostile attitude towards
preprints, mostly due to the lack of clear policy of the publishers. As an
example, Wiley-Blackwell, which publishes some of the leading journal in the
field, has no official policy on the subject \ref{table:policies}.

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in biology. Some publishers
tolerate preprints except for a few of their medical journals, e.g.: Journal
of the National Cancer Institute from Oxford and The Lancer from Elsevier.}
    \label{table:policies}
\end{table*}

There are other ``cultural'' barriers to adoption of preprint servers by
biologists.  The most notable preprint server,
arXiv, is designed to work best with \TeX{}, a nearly universal document
preparation system among physicists and mathematicians.  Jackson
\cite{jackson2002preprints} further argues that \TeX{} introduced an ``open
source mindset'' to its users, who now freely share their research findings as
well as their software.  Many biologists instead use proprietary software to
prepare their research findings, and many journals officially prefer Microsoft
Word documents as submissions.  The recent discipline-wide adoption of free
software packages such as R and associated document generation systems
\cite{xie2012}, as well as interest in open access publishing, has coincided
with recent rise in use of preprint servers by biologists, supporting Jackson's
claims. \cite{xie12}

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in mathematics and most branches of physics.
Physicist Paul Ginsparg originated arXiv in 1991 for theoretical high-energy
physicists to communicate preprints via email and ftp, and soon thereafter
adopted the newly created world-wide web\cite{jackson2002preprints}.  arXiv now
receives over 7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}).  arXiv divides its
submissions into subcategories of physics, mathematics, computer science,
quantitative biology, finance and statistics.  The quantitative biology category
includes subcategories for Populations and Evolution, Quantitative Methods and
other categories that may be of interest to biologists.

@@

%The average time to publication vary a lot between journal and can be as low as
%104 days (Evolution for 2011) to 213 (PLOS One in 2010)
